---
title: 'Project 2 : Cloud Data '
author: "Richard Fremgen (Duke ID: 1078591), Matthew Dockman (Duke ID: 0738472)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r setup, message=F, warning=F, echo=F}
library(tidymodels)
library(parsnip)
library(tune)
library(discrim)
library(ggpubr)
library(class)
library(caret)
library(tidyverse)
require(magrittr)
library(GGally)
require(plyr)
library(corrplot)
library(knitr)
library(class)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
```


***

## 0 - Load in Data

```{r, echo=FALSE}
# Load in three image data sets and merge into one data frame 

# Vector for column names
m_names <- c("y", "x", "label", "ndai", "sd", "corr", "ra_df", "ra_cf", "ra_bf", "ra_af", "ra_an") 

# M1 Data 
m1 <- as.data.frame(read.table("image_data/imagem1.txt",
                               col.names = m_names)) %>% 
                      mutate(image = "M1 Image",
                             expert_label = case_when(.$label == 1 ~ "Cloud",
                                                      .$label == 0 ~ "Unlabeled",
                                                      .$label == -1 ~ "Not Cloud"),
                             label = as.factor(label))
m1$expert_label <- as.factor(m1$expert_label)

# M2 Data 
m2 <- as.data.frame(read.table("image_data/imagem2.txt",
                               col.names = m_names)) %>% 
                      mutate(image = "M2 Image",
                             expert_label = case_when(.$label == 1 ~ "Cloud",
                                                      .$label == 0 ~ "Unlabeled",
                                                      .$label == -1 ~ "Not Cloud"),
                             label = as.factor(label))
m2$expert_label <- as.factor(m2$expert_label)

# M3 Data 
m3 <- as.data.frame(read.table("image_data/imagem3.txt",
                               col.names = m_names)) %>% 
                      mutate(image = "M3 Image",
                             expert_label = case_when(.$label == 1 ~ "Cloud",
                                                      .$label == 0 ~ "Unlabeled",
                                                      .$label == -1 ~ "Not Cloud"),
                             label = as.factor(label))
m3$expert_label <- as.factor(m3$expert_label)
 
# Combine three image files
m_all <- rbind(m1, m2, m3)

# Create a filtered data set - remove unlabeled points
m_filtered <- m_all %>%
  filter(expert_label != "Unlabeled")

# Are there any null values? 
colSums(is.na(m_all)) # No
```


## 1 - Data Collection and Exploration 

 
### 1B - Plot Maps using x,y coordinates with expersts labels are color of regions


```{r, fig.height = 3, echo=FALSE}
# M1 Data 
m1_plot <- ggplot(data = m1, aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue", "grey47")) +
  ggtitle("M1 Image") +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.text = element_text(size = 10),
        legend.title = element_blank())
labs(color = "Expert Label") 

# M2 Data 
m2_plot <- ggplot(data = m2, aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue", "grey47")) +
  ggtitle("M2 Image") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.title = element_blank()) +
  labs(color = "Expert Label")

# M3 Data 
m3_plot <- ggplot(data = m3, aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue", "grey47")) +
  ggtitle("M3 Image") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.title = element_blank()) +
  labs(color = "Expert Label")

# combine three plots into one
ggarrange(m1_plot, m2_plot, m3_plot, nrow = 1, common.legend = TRUE, legend ="bottom")
```

### 1B - Pictures with unlabeled dropped

```{r, eval=FALSE, echo=FALSE}
# M1 Data 
m1_plot <- m_filtered %>% filter(image == "M1 Image") %>%
  ggplot(aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue")) +
  ggtitle("M1 Image") +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.text = element_text(size = 10),
        legend.title = element_blank())
labs(color = "Expert Label") 

# M2 Data 
m2_plot <- m_filtered %>% filter(image == "M2 Image") %>%
  ggplot(aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue")) +
  ggtitle("M2 Image") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.title = element_blank()) +
  labs(color = "Expert Label")

# M3 Data 
m3_plot <- m_filtered %>% filter(image == "M3 Image") %>%
  ggplot(aes(x=x, y=y)) +
  geom_point(aes(color = expert_label)) +
  scale_color_manual(values = c("azure2", "skyblue")) +
  ggtitle("M3 Image") +
  theme(plot.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 10),
        legend.title = element_blank()) +
  labs(color = "Expert Label")

# combine three plots into one
ggarrange(m1_plot, m2_plot, m3_plot, nrow = 1, common.legend = TRUE, legend ="bottom")

```


### 1B - Sumarize the data (% of pixels for the differnt classes)

USE IN REPORT

```{r, echo=FALSE}
# Create a % table by image 
tab <- table(m_all$image, m_all$expert_label) 
tab[] <- sprintf("%.1f%%",round(prop.table(tab,margin = 1)*100, 3))

# Create a % table for all images combines
Total <- table(m_all$expert_label)
Total[] <- sprintf("%.1f%%",round(prop.table(Total)*100, 3))

# Combine two % tables into one 
rbind(tab, Total) %>% 
  kable(align = "c",
        caption = "% of Pixels by Different Expert Classes")

# Stop knit here
#knitr:::knit_exit()
```

### 1B - Sumarize the data (% of pixels for the differnt classes)

If you want to look at sd, corr, ndai on an image level 

```{r}
# All Data Summary Table 

print1 <- m_all %>% 
  group_by(expert_label) %>%
  dplyr::summarise(count = n(),
                   ndai = round(mean(ndai),3),
                   sd = round(mean(sd),3),
                   corr = round(mean(corr),3),
                   ra_df = round(mean(ra_df),3),
                   ra_cf = round(mean(ra_cf),3),
                   ra_bf = round(mean(ra_bf),3),
                   ra_af = round(mean(ra_af),3),
                   ra_an = round(mean(ra_an),3))
print1 %>%
  kable(align = "c")


# Stop knit here
knitr:::knit_exit()

```


```{r, echo=FALSE, eval=FALSE}

# All Data Summary Table 
m_all %>% 
  group_by(expert_label) %>%
  dplyr::summarise(count = n(),
                   percent = count / nrow(m_all),
                   avg_ndai = mean(ndai),
                   avg_sd = mean(sd),
                   avg_corr = mean(corr)) 

# M1 Image 
m1 %>% 
  group_by(expert_label) %>%
  dplyr::summarise(count = n(),
                   percent = count / nrow(m2),
                   avg_ndai = mean(ndai),
                   avg_sd = mean(sd),
                   avg_corr = mean(corr))

# M2 Image
m2 %>% 
  group_by(expert_label) %>%
  dplyr::summarise(count = n(),
                   percent = count / nrow(m2),
                   avg_ndai = mean(ndai),
                   avg_sd = mean(sd),
                   avg_corr = mean(corr)) 

# M3 Image 
m3 %>% 
  group_by(expert_label) %>%
  dplyr::summarise(count = n(),
                   percent = count / nrow(m3),
                   avg_ndai = mean(ndai),
                   avg_sd = mean(sd),
                   avg_corr = mean(corr))
```

### 1C - Pairwise Scatterplots  

NEW PLOTS

```{r, echo=FALSE, eval=FALSE} 
# knitr::knit_exit()

# Specify colors
my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")  

# All Images 
m_all %>%
  select(4:6) %>%
  ggpairs(aes(color = m_all$expert_label),
          title = "All Images", 
          diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
          lower = list(continuous = wrap("points", alpha = 0.05, size=0.1)),
          legend = 1) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))

# M1 Image 
m1 %>%
  select(4:6) %>%
  ggpairs(aes(color = m1$expert_label),
          title = "M1 Image", 
          diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
          lower = list(continuous = wrap("points", alpha = 0.05, size=0.1)),
          legend = 1) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))

# M2 Image 
m2 %>%
  select(4:6) %>%
  ggpairs(aes(color = m2$expert_label),
          title = "M2 Image", 
          diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
          lower = list(continuous = wrap("points", alpha = 0.05, size=0.1)),
          legend = 1) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))

# M3 Image
m3 %>%
  select(4:6) %>%
  ggpairs(aes(color = m3$expert_label),
          title = "M3 Image", 
          diag = list(continuous = wrap("densityDiag", alpha = 0.5)),
          lower = list(continuous = wrap("points", alpha = 0.05, size=0.1)),
          legend = 1) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 10),
        plot.title = element_text(hjust = 0.5))
```

### 1C - Pairwise Scatterplots with adjusting CEX value and coloring by label - (ALL Data Combnied)

OLD PLOTS

```{r, eval=FALSE, echo=FALSE}
# Specify colors
my_cols <- c("#00AFBB", "#E7B800", "#FC4E07")  

# All Data Combined 
pairs(m_all[4:6],
      cex = 0.05,
      col = my_cols[m_all$label],
      main = "All Data (0.05)")

pairs(m_all[4:6],
      cex = 0.01,
      col = my_cols[m_all$label],
      main = "All Data (0.01)")

# M1 Image 
 pairs(m1[4:6],
      cex = 0.05,
      col = my_cols[m_all$label],
      main = "M1 Image Colored (0.05)")

# M2 Image 
pairs(m2[4:6],
      cex = 0.05,
      col = my_cols[m_all$label],
      main = "M2 Image  Colored (0.05)")

# M3 Image 
pairs(m3[4:6],
      cex = 0.05,
      col = my_cols[m_all$label],
      main = "M3 Image Colored (0.05)")
```

### 1C - Correlation Plots - ALL DATA POINTS

```{r, eval=FALSE, echo=FALSE}
# All Data Combined
all_cor <- cor(m_all[4:11])
corrplot(all_cor, method = "number", type = "upper", title = "All Data", mar=c(0,0,2,0))
corrplot(all_cor, method = "square", type = "upper", title = "All Data", mar=c(0,0,2,0))

# M1 Image
m1_cor <- cor(m1[4:11])
corrplot(m1_cor, method = "number", type = "upper", title = "M1 Image", mar=c(0,0,2,0))
corrplot(m1_cor, method = "square", type = "upper", title = "M1 Image", mar=c(0,0,2,0))

# M2 Image Data Combined
m2_cor <- cor(m2[4:11])
corrplot(m2_cor, method = "number", type = "upper", title = "M2 Image", mar=c(0,0,2,0))
corrplot(m2_cor, method = "square", type = "upper", title = "M2 Image", mar=c(0,0,2,0)) 

# M3 Image Data Combined
m3_cor <- cor(m3[4:11])
corrplot(m3_cor , method = "number", type = "upper", title = "M3 Image", mar=c(0,0,2,0))
corrplot(m3_cor , method = "square", type = "upper", title = "M3 Image", mar=c(0,0,2,0))
```

### 1C - Correlation Plots - UNLABELED REMOVED 

These plots allow us to look at the correlation between exper_label and the features of interest

```{r, echo=FALSE, eval=FALSE}
# All Data Combined
m_fil_cor <- m_filtered %>%
  mutate(label = ifelse(label == -1, 0, 1))

all_cor2 <- m_fil_cor %>%
  select(4:11, label) %>%
  cor() 

corrplot(all_cor2, method = "number", type = "upper", title = "All Data", mar=c(0,0,2,0))
corrplot(all_cor2, method = "square", type = "upper", title = "All Data", mar=c(0,0,2,0))

# M1 Image
m1_cor2 <- m_fil_cor %>%
  filter(image == "M1 Image") %>%
  select(4:11, label) %>%
  cor() 

corrplot(m1_cor2, method = "number", type = "upper", title = "M1 Image", mar=c(0,0,2,0))
corrplot(m1_cor2, method = "square", type = "upper", title = "M1 Image", mar=c(0,0,2,0))

# M2 Image Data Combined
m2_cor2 <- m_fil_cor %>%
  filter(image == "M2 Image") %>%
  select(4:11, label) %>%
  cor() 

corrplot(m2_cor2, method = "number", type = "upper", title = "M2 Image", mar=c(0,0,2,0))
corrplot(m2_cor2, method = "square", type = "upper", title = "M2 Image", mar=c(0,0,2,0)) 

# M3 Image Data Combined
m3_cor2 <- m_fil_cor %>%
  filter(image == "M3 Image") %>%
  select(4:11, label) %>%
  cor() 

corrplot(m3_cor2 , method = "number", type = "upper", title = "M3 Image", mar=c(0,0,2,0))
corrplot(m3_cor2 , method = "square", type = "upper", title = "M3 Image", mar=c(0,0,2,0))
```


### 1C - Grid of Variable Distribution - ALL EXPERT LABELS 

Run entire code chunk at once

```{r, echo=FALSE, eval=FALSE}
p1 <- ggplot(data = m_all, aes(x = ndai)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p2 <- ggplot(data = m_all, aes(x = sd)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p3 <- ggplot(data = m_all, aes(x = corr)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p4 <- ggplot(data = m_all, aes(x = ra_df)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p5 <- ggplot(data = m_all, aes(x = ra_cf)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p6 <- ggplot(data = m_all, aes(x = ra_bf)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p7 <- ggplot(data = m_all, aes(x = ra_af)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p8 <- ggplot(data = m_all, aes(x = ra_an)) + geom_density(aes(fill = expert_label), alpha = 0.5)
ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, nrow = 3, ncol = 3, common.legend = TRUE, legend ="bottom") %>%
  annotate_figure(top = "Feature Density Comparison (All Expert Labels)")
```

### 1C - Grid of Variable Distribution - UNLABELED REMOVED 

Run entire code chunk at once

```{r, echo=FALSE}
p1 <- ggplot(data = m_filtered, aes(x = ndai)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p2 <- ggplot(data = m_filtered, aes(x = sd)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p3 <- ggplot(data = m_filtered, aes(x = corr)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p4 <- ggplot(data = m_filtered, aes(x = ra_df)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p5 <- ggplot(data = m_filtered, aes(x = ra_cf)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p6 <- ggplot(data = m_filtered, aes(x = ra_bf)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p7 <- ggplot(data = m_filtered, aes(x = ra_af)) + geom_density(aes(fill = expert_label), alpha = 0.5)
p8 <- ggplot(data = m_filtered, aes(x = ra_an)) + geom_density(aes(fill = expert_label), alpha = 0.5)
ggarrange(p1, p2, p3, p4, p5, p6, p7, p8, nrow = 3, ncol = 3, common.legend = TRUE, legend ="bottom") %>%
  annotate_figure(top = "Feature Density Comparison (Unlabled Removed)")
```

### 1C - BOXPLOT COMPARISON - ALL EXPERT LABELS 

Run entire code chunk at once

```{r, eval=FALSE, echo=FALSE}
# SD, CORR< NDAI, RA_DF Boxplots
p1 <- ggplot(data = m_all, aes(x = expert_label, y = ndai)) + geom_boxplot(alpha = 0.1)
p2 <- ggplot(data = m_all, aes(x = expert_label, y = sd)) + geom_boxplot(alpha = 0.1)
p3 <- ggplot(data = m_all, aes(x = expert_label, y = corr)) + geom_boxplot(alpha = 0.1)
p4 <- ggplot(data = m_all, aes(x = expert_label, y = ra_df)) + geom_boxplot(alpha = 0.1)
ggarrange(p1, p2, p3, p4, nrow = 2, ncol = 2) %>%
  annotate_figure(top = "Feature Boxplot Comparison I (All Expert Labels)")

# Radiance Angle Boxplots 
p5 <- ggplot(data = m_all, aes(x = expert_label, y = ra_cf)) + geom_boxplot(alpha = 0.1)
p6 <- ggplot(data = m_all, aes(x = expert_label, y = ra_bf)) + geom_boxplot(alpha = 0.1)
p7 <- ggplot(data = m_all, aes(x = expert_label, y = ra_af)) + geom_boxplot(alpha = 0.1)
p8 <- ggplot(data = m_all, aes(x = expert_label, y = ra_an)) + geom_boxplot(alpha = 0.1)
ggarrange(p5, p6, p7, p8, nrow = 2, ncol = 2) %>%
  annotate_figure(top = "Feature Boxplot Comparison II (All Expert Labels)")
```

### 1C - Histograms of Different Paramters

```{r, echo=FALSE, eval=FALSE}
# NDAI - All Data 
ggplot(data = m_all, aes(x = ndai)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("All Data - NDAI") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~expert_label)

# NDAI - By Image 
ggplot(data = m_all, aes(x = ndai)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("NDAI By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label) 

# CORR - All Data 
ggplot(data = m_all, aes(x = corr)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("All Data - CORR") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~expert_label)

# CORR - By Image 
ggplot(data = m_all, aes(x = corr)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("CORR By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label) 

# SD - All Data 
ggplot(data = m_all, aes(x = sd)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("All Data - SD") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~expert_label)

# SD - By Image 
ggplot(data = m_all, aes(x = sd)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("SD By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)

# RA_DF
ggplot(data = m_all, aes(x = ra_df)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("RA_DF By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)

# RA_CF
ggplot(data = m_all, aes(x = ra_cf)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("RA_CF By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)

# RA_BF
ggplot(data = m_all, aes(x = ra_bf)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("RA_BF By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)

# RA_AF
ggplot(data = m_all, aes(x = ra_af)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("RA_AF By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)

# RA_AN
ggplot(data = m_all, aes(x = ra_an)) +
  geom_histogram(bins = 20, color = "grey") +
  ggtitle("RA_AN By Image") +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_grid(image~expert_label)
```

PCA on M1

```{r, eval=FALSE, echo=FALSE}
m1_pca <- m1 %>% 
  filter(expert_label != "Unlabeled")
m1_use <- m1_pca %>%
  select(4:11) 


m1_df <- as.data.frame(prcomp(m1_use, scale = TRUE)$x) %>%
  mutate(expert_label = m1_pca$expert_label)

# plot pca
ggplot(m1_df, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = m1_df$expert_label), alpha = 0.2)
```


PCA on M2


```{r, echo=FALSE, eval=FALSE}
m3_pca <- m3 %>% 
  filter(expert_label != "Unlabeled")

m3_use <- m3_pca %>%
  select(4:11) 

m3_pca2 <- prcomp(m3_use, scale = TRUE)

m3_df <- as.data.frame(m3_pca2$x) %>%
  mutate(expert_label = m3_pca$expert_label)

# plot pca
ggplot(m3_df, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = m3_df$expert_label), alpha = 0.2)

# extract loadings
m3_pca2$rotation[,1:2]
(m3_pca2$sdev)^2/(sum((m3_pca2$sdev)^2)) 

```

PCA on all data

```{r,eval=FALSE, echo=FALSE}
mall_pca <- m_all %>% 
  filter(expert_label != "Unlabeled")

mall_use <- mall_pca %>%
  select(4:11) 

mall_pca2 <- prcomp(mall_use, scale = TRUE)

mall_df <- as.data.frame(mall_pca2$x) %>%
  mutate(expert_label = mall_pca$expert_label)

# plot pca
ggplot(mall_df, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = mall_df$expert_label), alpha = 0.2)

# extract loadings
mall_pca2$rotation[,1:2]
(mall_pca2$sdev)^2/(sum((mall_pca2$sdev)^2)) 
```

## 2 - Preparation

### 2A - Splitting the entire data set 

DATA SEPARATION 1 - Write a function to divide data into four equal boxes (BLOCK)

```{r, echo=FALSE}
# Function will take into an image data frame and divide the pixels into four equals boxes
image.block <- function(df) {
  df<- df %>%
  dplyr::mutate(box_num_y = cut(df$y, breaks = 4, labels = c("box1", "box2", "box3", "box4")),
         box_num_x = cut(df$x, breaks = 4, labels = c("box1", "box2", "box3", "box4")),
         box_num = case_when(box_num_y == "box1" & box_num_x == "box1" ~ "box1",
                             box_num_y == "box2" & box_num_x == "box1" ~ "box2",
                             box_num_y == "box3" & box_num_x == "box1" ~ "box3",
                             box_num_y == "box4" & box_num_x == "box1" ~ "box4",
                             box_num_y == "box1" & box_num_x == "box2" ~ "box5",
                             box_num_y == "box2" & box_num_x == "box2" ~ "box6",
                             box_num_y == "box3" & box_num_x == "box2" ~ "box7",
                             box_num_y == "box4" & box_num_x == "box2" ~ "box8",
                             box_num_y == "box1" & box_num_x == "box3" ~ "box9",
                             box_num_y == "box2" & box_num_x == "box3" ~ "box10",
                             box_num_y == "box3" & box_num_x == "box3" ~ "box11",
                             box_num_y == "box4" & box_num_x == "box3" ~ "box12",
                             box_num_y == "box1" & box_num_x == "box4" ~ "box13",
                             box_num_y == "box2" & box_num_x == "box4" ~ "box14",
                             box_num_y == "box3" & box_num_x == "box4" ~ "box15",
                             box_num_y == "box4" & box_num_x == "box4" ~ "box16"
                             ),
         box_num_y = NULL,
         box_num_x = NULL)
  return(df) 
}

# Run image.block on all three images 
m1<- image.block(m1) 
m2 <- image.block(m2) 
m3 <- image.block(m3)

m1 <- m1 %>%
  mutate(block = case_when(box_num == "box3" | box_num == "box4" | box_num == "box7" | box_num == "box8" ~ 1,
                           box_num == "box1" | box_num == "box12" | box_num == "box15" | box_num == "box16" ~ 2,
                           box_num == "box2" | box_num == "box5" | box_num == "box9" | box_num == "box13" ~ 3,
                           box_num == "box6" | box_num == "box10" | box_num == "box11" | box_num == "box14" ~ 4)) %>% filter(label != 0)

m2 <- m2 %>%
  mutate(block = case_when(box_num == "box1" | box_num == "box3" | box_num == "box8" | box_num == "box16" ~ 5,
                           box_num == "box4" | box_num == "box11" | box_num == "box12" | box_num == "box14" ~ 6,
                           box_num == "box2" | box_num == "box6" | box_num == "box7" | box_num == "box10" ~ 7,
                           box_num == "box5" | box_num == "box9" | box_num == "box13" | box_num == "box15" ~ 8)) %>% filter(label != 0)

m3 <- m3 %>%
  mutate(block = case_when(box_num == "box1" | box_num == "box4" | box_num == "box13" | box_num == "box16" ~ 9,
                           box_num == "box6" | box_num == "box8" | box_num == "box10" | box_num == "box12" ~ 10,
                           box_num == "box3" | box_num == "box7" | box_num == "box11" | box_num == "box15" ~ 11,
                           box_num == "box2" | box_num == "box5" | box_num == "box9" | box_num == "box14" ~ 12)) %>% filter(label != 0)

# Combine and update the m_all data frame 
# m_all <- rbind(m1, m2, m3)
```


DATA SEPARATION 2 -  Use kmeans clustering to split data - second way to split data (CLUSTER)


```{r, echo=FALSE}
# M1 Image 
set.seed(127)
m1_k <- kmeans(m1[,1:2], centers = 4)
m1 <- m1 %>% 
  mutate(cl_num = m1_k$cluster,
         cluster = cl_num,
         cl_num = NULL)

# M2 Image 
set.seed(127)
m2_k <- kmeans(m2[,1:2], centers = 4)
m2 <- m2 %>% 
  mutate(cl_num = m2_k$cluster,
         cluster = case_when(cl_num == 1 ~ 9,
                             cl_num == 2 ~ 10,
                             cl_num == 3 ~ 11,
                             cl_num == 4 ~ 12),
         cl_num = NULL)

# M3 Image 
set.seed(127)
m3_k <- kmeans(m3[,1:2], centers = 4)
m3 <- m3 %>% 
  mutate(cl_num = m3_k$cluster,
         cluster = case_when(cl_num == 1 ~ 5,
                             cl_num == 2 ~ 6,
                             cl_num == 3 ~ 7,
                             cl_num == 4 ~ 8),
         cl_num = NULL)

# Combine and update the m_all data frame 
m_all <- rbind(m1, m2, m3)
```


DATA SEPARATION 1  - Now separate into training, validation, and testing sets (BLOCK) 


```{r, echo=FALSE}
# Now Split into Three sets: Training, Testing, Validation 

# Testing Set 
set.seed(127)
# set.seed(35)
block_test <- sample(unique(m_all$block), size = 2)
m_test <- m_all %>%
  filter(block %in% block_test) 

# Training/Val Set Combined 
m_train_val <- m_all %>%
  filter(!(block %in% block_test))

# If we want to separate training and validation blocks 
set.seed(127)
# set.seed(35)
block_val <- sample(unique(m_train_val$block), size = 1) 
m_val <- m_train_val %>%
  filter(block %in% block_val) 
m_train <- m_train_val %>%
  filter(!(block %in% block_val))
```


DATA SEPARATION 2  - Now separate into training, validation, and testing sets (CLUSTER)


```{r, echo=FALSE}
# Now Split into Three sets: Training, Testing, Validation 

# Testing Set 
set.seed(127)

cluster_test <- sample(unique(m_all$cluster), size = 2)
m_test2 <- m_all %>%
  filter(cluster %in% cluster_test) 

# Training/Val Set Combined 
m_train_val2 <- m_all %>%
  filter(!(cluster %in% cluster_test))

# If we want to separate training and validation blocks 
set.seed(127)
cluster_val <- sample(unique(m_train_val2$cluster), size = 1) 
m_val2 <- m_train_val2 %>%
  filter(cluster %in% cluster_val) 
m_train2 <- m_train_val2 %>%
  filter(!(cluster %in% cluster_val))
```


Testing to ensure blocking functions worked:


```{r, eval=FALSE, echo=FALSE}
# Visualize data to verify 
ggplot(data = m1, aes(x=x, y=y)) +
  geom_point(aes(color = box_num)) +
  ggtitle("Box Test Viz") +
  theme(plot.title = element_text(hjust = 0.5), 
        legend.text = element_text(size = 10),
        legend.title = element_blank())

# Equal proportions? 
m2 %>% group_by(box_num) %>% dplyr::summarise(c = n())
m_all %>% group_by(block) %>% dplyr::summarise(c = n()) 
```

### 2C - Trivial Classifier

DATA SEPARATION 1  -  (BLOCK) 

What would be the accuracy of the model if we set all observations in the test set to -1?

```{r, echo=FALSE}
tbl1 <- m_all %>% 
  group_by(expert_label) %>%
  dplyr::summarise(n = n(),
                   Percent = round(n /nrow(m_all) * 100, 1)) %>%
  mutate(Type = "All Data") 

# Training Set 

tbl2a <- m_train %>% 
  mutate(Type = "Training Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_train) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_train %>% 
  mutate(Type = "Training Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_train) * 100, 1),
                   n = nrow(m_train)) %>%
  dplyr::select(-c("num", "Type"))

tbl_train <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Validation Set 

tbl2a <- m_val %>% 
  mutate(Type = "Validation Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_val) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_val %>% 
  mutate(Type = "Validation Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_val) * 100, 1),
                   n = nrow(m_val)) %>%
  dplyr::select(-c("num", "Type"))

tbl_val <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Testing Set 

tbl2a <- m_test %>% 
  mutate(Type = "Testing Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_test) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_test %>% 
  mutate(Type = "Testing Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_test) * 100, 1),
                   n = nrow(m_test)) %>%
  dplyr::select(-c("num", "Type"))

tbl_test <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# All Data

tbl2a <- m_all %>% 
  mutate(Type = "All Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_all) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_all %>% 
  mutate(Type = "All Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_all) * 100, 1),
                   n = nrow(m_all)) %>%
  dplyr::select(-c("num", "Type"))

tbl_all <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Combine Tables Together
tbl_combine <- rbind(tbl_train, tbl_val, tbl_test, tbl_all) 
tbl_combine

# Combine two % tables into one 
tbl_combine %>%
  kable(align = "c",
        caption = "Block Splitting: Expert Labels by Data Sets")
```

### 2C - Trivial Classifier

DATA SEPARATION 2  -  (CLUSTER) 

What would be the accuracy of the model if we set all observations in the test set to -1?

```{r, echo=FALSE}
tbl1 <- m_all %>% 
  group_by(expert_label) %>%
  dplyr::summarise(n = n(),
                   Percent = round(n /nrow(m_all) * 100, 1)) %>%
  mutate(Type = "All Data") 

# Training Set 

tbl2a <- m_train2 %>% 
  mutate(Type = "Training Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_train2) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_train2 %>% 
  mutate(Type = "Training Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_train2) * 100, 1),
                   n = nrow(m_train2)) %>%
  dplyr::select(-c("num", "Type"))

tbl_train <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Validation Set 

tbl2a <- m_val2 %>% 
  mutate(Type = "Validation Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_val2) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_val2 %>% 
  mutate(Type = "Validation Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_val2) * 100, 1),
                   n = nrow(m_val2)) %>%
  dplyr::select(-c("num", "Type"))

tbl_val <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Testing Set 

tbl2a <- m_test2 %>% 
  mutate(Type = "Testing Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_test2) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_test2 %>% 
  mutate(Type = "Testing Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_test2) * 100, 1),
                   n = nrow(m_test2)) %>%
  dplyr::select(-c("num", "Type"))

tbl_test <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# All Data

tbl2a <- m_all %>% 
  mutate(Type = "All Data") %>%
  group_by(Type) %>%
  filter(label == 1) %>%
  dplyr::summarise(num = n(),
                   Cloud = round(num / nrow(m_all) * 100, 1)) %>%
  dplyr::select(-c("num"))

tbl2b <- m_all %>% 
  mutate(Type = "All Data") %>%
  group_by(Type) %>%
  filter(label == -1) %>%
  dplyr::summarise(num = n(),
                   nc = round(num / nrow(m_all) * 100, 1),
                   n = nrow(m_all)) %>%
  dplyr::select(-c("num", "Type"))

tbl_all <- cbind(tbl2a, tbl2b) %>%
  mutate(Cloud = sprintf("%.1f%%",Cloud),
         nc = sprintf("%.1f%%",nc),
         "Not Cloud" = nc) %>%
  dplyr::select(c("Type", "Cloud", "Not Cloud", "n"))

# Combine Tables Together
tbl_combine <- rbind(tbl_train, tbl_val, tbl_test, tbl_all) 
tbl_combine

# Combine two % tables into one 
tbl_combine %>%
  kable(align = "c",
        caption = "Cluster Splitting: Expert Labels by Data Sets")
```



***

